---
title: "HW9"
author: "Natalie Brewer"
date: "2023-11-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyr)
library(ggplot2)
```

## Problem 9A (12.26)

```{r}
dogs_data <- read.table("/Users/nataliebrewer/Desktop/Stat 135/HW9/dogs.txt", sep=" ")
dogs_df <- data.frame(dogs_data)
dogs_df

T1 <- unname(unlist(dogs_df[1,]))
T2 <- unname(unlist(dogs_df[2,]))
T3 <- unname(unlist(dogs_df[3,]))

groups <- c(rep("T1",10),rep("T2",10),rep("T3",10))
treatments <- c(T1,T2,T3)
dogs_tidy <- data.frame(groups,treatments)
head(dogs_tidy)

dogs_tidy %>% ggplot(aes(x=groups,y=treatments)) + geom_boxplot()
qqnorm(T1)
qqnorm(T2)
qqnorm(T3)
```

1) F Test

```{r}
# Worked out by hand
I=3
J=10

mean_T1 <- mean(T1)
mean_T2 <- mean(T2)
mean_T3 <- mean(T3)
c(mean_T1, mean_T2, mean_T3)

mean_tot <- mean(treatments)
mean_tot

var_T1 <- var(T1)
var_T2 <- var(T2)
var_T3 <- var(T3)
c(var_T1, var_T2, var_T3)

SSW <- (J-1) * (var_T1 + var_T2 + var_T3)
SSW
SSB <- J * ((mean_T1 - mean_tot)^2 + (mean_T2 - mean_tot)^2 + (mean_T3 - mean_tot)^2)
SSB

F <- (SSB/(I-1)) / (SSW/(I*(J-1)))
F

1 - pf(F,2,27)

# Check answer with R
oneway.test(treatments~groups,var.equal = TRUE)
```
2) Bonferroni Test

```{r}
# Just use R
pairwise.t.test(dogs_tidy$treatments,dogs_tidy$groups,p.adjust="none" )
```
We compare these value with alpha/3 = 0.017

Based on the F-test and the Bonferroni test, there appear to be differences between the effects of the three treatments. Specifically, there is a difference between T3, which is Cyclopropane, and the other two groups.

3) Non-parametric: Kruskal Wallis Test

```{r}
# Just use R
kruskal.test(treatments~groups)
```

Based on the Kruskal-Wallis test, we would fail to reject the null hypothesis if using a significance level lower that 0.05. Otherwise, we would conclude that there is no difference between the treatments.

## Problem 9B (12.27)
```{r}
AJ <- unname(unlist(read.table("/Users/nataliebrewer/Desktop/Stat 135/HW9/aj.txt", sep=" ")))
C57 <- unname(unlist(read.table("/Users/nataliebrewer/Desktop/Stat 135/HW9/c57.txt", sep=" ")))
F2 <- unname(unlist(read.table("/Users/nataliebrewer/Desktop/Stat 135/HW9/f2.txt", sep=" ")))

mice_treatments <- c(AJ, C57, F2)
mice_groups <- c(rep("AJ",length(AJ)), rep("C57",length(C57)), rep("F2",length(F2)) )

mice_tidy <- data.frame(mice_groups,mice_treatments)
head(mice_tidy)

# Boxplots
mice_tidy %>% ggplot(aes(x=mice_groups,y=mice_treatments)) + geom_boxplot()

# Bonferroni test
pairwise.t.test(mice_tidy$mice_treatments,mice_tidy$mice_groups,p.adjust="bonferroni" )

```

Since we used the bonferroni adjusment for the p-values, we compare these results with the alpha level 0.05.

Because the p values are all below 0.05, we reject the null hypothesis and conclude that there are significant differences between the species of mice.

## Problem 9C (12.28)
```{r}

Typ1 <- c(1.7, 1.9, 6.1, 12.5, 16.5, 25.1, 30.5, 42.1, 82.5)
Typ2 <- c(13.6, 19.8, 25.2, 46.2, 46.2, 61.1)
Typ3 <- c(13.4, 20.9, 25.1, 29.7, 46.9)

watch_results <- c(Typ1, Typ2, Typ3)
watch_types <- c(rep("Typ1",length(Typ1)), rep("Typ2",length(Typ2)), rep("Typ3",length(Typ3)) )

watch_tidy <- data.frame(watch_types,watch_results)
head(watch_tidy)

# Boxplots
watch_tidy %>% ggplot(aes(x=watch_types,y=watch_results)) + geom_boxplot()

# F test
oneway.test(watch_results~watch_types,var.equal = TRUE)

# Bonferroni test
pairwise.t.test(watch_tidy$watch_results,watch_tidy$watch_types,p.adjust="bonferroni")

# Kruskal Wallis test
kruskal.test(watch_results~watch_types)
```

Based on the results of both the parametric and nonparametric tests, we fail to reject the null hypothesis and conclude that the watch types are not significantly different.
