---
title: "HW3"
author: "Natalie Brewer"
date: "2023-09-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

## Problem 3F

```{r 3F}
digits = c(3,0,3,6,3,7,9,0,7,2)
SID <- data.frame(digits)

ggplot(SID, aes(x=digits)) +
  geom_histogram(aes(y=after_stat(density)), binwidth = 1) +
  labs(title="Density of Student ID Numbers", x="Digit", y="Density") +
  scale_x_continuous(breaks=seq(min(digits), max(digits), by=1)) + 
  theme_minimal()

#Find mean
mean <- mean(digits)
mean
#Find standard deviation
sd <- sd(digits)
sd
```

## Problem 3G
```{r 3G}
sample <- sample(digits, 400, replace=TRUE)
sample

sampleMean <- mean(sample)
sampleMean

standardError <- sd/sqrt(400)
standardError
```
The mean of the sampling distribution is the population mean, 4.

The standard error is $$\frac{\sqrt{\sigma^2}}{\sqrt{n}} = \frac{3.0912}{\sqrt{400}} = 0.155$$

```{r}
x_vals <- seq(3.5, 4.5, by=0.001)
df <- data.frame(x=x_vals, y=dnorm(x_vals, mean=mean, sd=standardError))

ggplot(df, aes(x=x, y=y)) +
  geom_line(color="blue", linewidth=1.2) +
  labs(title="Sampling Distribution", x="x", y="Density") +
  theme_minimal()
```

## Problem 3H
```{r 3H}
find_avg <- function(){
  resample <- sample(digits, size = 400, replace = T)
  xbar <- mean(resample)
  xbar
}

sampleMeanVec <- replicate(100, find_avg())
sampleMeanVec

sampleMeanDf <- data.frame(sampleMeanVec)

ggplot(sampleMeanDf, aes(x=sampleMeanVec)) +
  geom_histogram(aes(y=after_stat(density)), binwidth = .05) +
  labs(title="Sample Means Histogram", x="Sample Mean", y="Count") +
  scale_x_continuous(breaks=seq(3.5, 4.5, by=.1)) + 
  stat_function(fun = dnorm, args = list(mean = mean, sd =standardError), color = "blue", linewidth = 1.5) +
  theme_minimal()
```

## Problem 3I
```{r 3I}
CIs <- data.frame(lower = c(), upper = c())

find_CI <- function(mean, width){
  z <- qnorm(1-((100-width)/200))
  
  lower <- mean - (z*standardError)
  upper <- mean + (z*standardError)
  
  return(list(lower = lower, upper = upper))
}

for (i in sampleMeanVec){
  CIs <- rbind(CIs, find_CI(i,95))
}

CIs
```

```{r}
count <- 0  #this will be the number of intervals containing the actual population mean 

for (i in 1:nrow(CIs)){
  lower <- CIs[i,1]
  upper <- CIs[i,2]
  if((mean > lower) && (mean < upper)){
    count <- count + 1
  }
}
count
```



## Problem 3J
The probability that each interval contains the actual parameter is 95%. This is a Bernouli trial with p = .95. When we take 100 samples and look at the confidence interval determined by each, the distribution is Binomial(100, .95).
E(G) = np = 95
SE(G) = np(1-p) = 4.75
```{r 3J}
df <- data.frame(x = 0:100, probabilities = dbinom(0:100, size = 100, prob = 0.95))

ggplot(df, aes(x = x, y = probabilities)) +
  geom_col(fill = "lightblue", color = "black") + 
  geom_vline(aes(xintercept = count), color = "red", linetype = "dashed", size = 1) +
  labs(title = "Binomial Distribution of G", x = "Intervals containing parameter", y = "Probability") +
  theme_minimal()
```
Based on my values for E(G) and SE(G), the placement of the vertical line from part 3I makes sense. It is close to the mean.



